Supplement to ``{\mytitle}''

### 2023-09-01 si/calibrateMod3.Rnw
Started work to make individual si sections seperate

\pdfminorversion=4

\newcommand\secTitleSpace{\hspace{3mm}}

<<packages, include=FALSE>>=
library(tidyverse)
library(pomp)
library(haitipkg)
library(spatPomp)
library(doParallel)
library(doRNG)

@@

<<set-opts, include=FALSE>>=
options(
  scipen = 2,
  help_type = "html",
  stringsAsFactors = FALSE,
  continue = "+  ",
  width = 70,
  useFancyQuotes = FALSE,
  reindent.spaces = 2,
  xtable.comment = FALSE
)
@

<<knitr-opts, include=FALSE, purl=FALSE>>=
library(knitr)
opts_knit$set(concordance=TRUE)
opts_chunk$set(
  progress=TRUE,prompt=TRUE,highlight=FALSE,
  tidy=TRUE,
  tidy.opts=list(
    keep.blank.line=FALSE
  ),
  comment="",
  warning=FALSE,
  message=FALSE,
  error=TRUE,
  echo=FALSE,
  cache=FALSE,
  strip.white=TRUE,
  results="markup",
  fig.path="figure/",
  fig.lp="fig:",
  fig.align="center",
  fig.show="asis",
  dev="pdf",
  dev.args=list(
    bg="transparent",
    pointsize=9
  )
)

myround <- function (x, digits = 1) {
  # taken from the broman package
  if (digits < 1)
    stop("This is intended for the case digits >= 1.")
  if (length(digits) > 1) {
    digits <- digits[1]
    warning("Using only digits[1]")
  }
  tmp <- sprintf(paste("%.", digits, "f", sep = ""), x)
  zero <- paste0("0.", paste(rep("0", digits), collapse = ""))
  tmp[tmp == paste0("-", zero)] <- zero
  tmp
}

@

%% copied from pnas-new.cls
\usepackage{amsmath,amsfonts,amssymb,graphicx}
\usepackage{pict2e}
\usepackage{curve2e}

%%%% parameters %%%%%%%%%%5
\newcommand\Wsat{W_{\mathrm{sat}}}
\newcommand\muIR{\mu_{IR}}
\newcommand\muEI{\mu_{EI}}
\newcommand\transmission{\beta}
\newcommand\seasAmplitude{a}
\newcommand\rainfallExponent{r}
\newcommand\muRS{\mu_{RS}}
\newcommand\vaccineEfficacy{\theta}
\newcommand\muBirth{\mu_S}
\newcommand\muDeath{\delta}
\newcommand\choleraDeath{\delta_{C}}
\newcommand\symptomFrac{f}
\newcommand\asymptomRelativeInfect{\epsilon}
\newcommand\asymptomRelativeShed{\epsilon_{W}}
\newcommand\Wbeta[1]{\beta_{W#1}}
\newcommand\Wremoval{\delta_W}
\newcommand\Wshed{\mu_W}
\newcommand\mixExponent{\nu}
\newcommand\sigmaProc{\sigma_{\mathrm{proc}}}
\newcommand\reportRate{\rho}
\newcommand\obsOverdispersion{\psi}
\newcommand\phaseParm{\phi}
\newcommand\transmissionTrend{\zeta}
\newcommand\Binit{\xi}
\newcommand\vaccClass{Z}
\newcommand\vaccCounter{z}
\newcommand\modelCounter{m}
\newcommand\Whur[1]{\beta_{W#1}^{hm}}
\newcommand\hHur[1]{h_{#1}^{hm}}
\newcommand\Iinit{I_{0,0}}
\newcommand\Einit{E_{0, 0}}
\newcommand\reportChange{\delta}
\newcommand\NBintercept{\alpha}
\newcommand\NBsize{\varphi}

\newcommand\missing{\hspace{6mm} ---}
\newcommand\fixed{\color{blue}}
\newcommand\demography{\bullet}

\newcommand\mytitle{Informing policy via dynamic models: Cholera in Haiti}
\newcommand\comma{{\hspace{-0.25mm},\hspace{-0.2mm}}}

%%%%%%%%%%%%%%%% START OF USER DEFINITIONS %%%%%%%%%%%%%%%5

\newcommand\siOnly[1]{} %% redefined for the supplement
\newcommand\msOnly[1]{#1} %% redefined for the supplement

\newcommand\code[1]{\texttt{#1}}
\usepackage{xurl}
% \usepackage[breaklinks]{hyperref}
\usepackage{makecell}
\usepackage{multirow}
\usepackage{multicol}
\usepackage[ruled,noline,linesnumbered]{algorithm2e}
% \usepackage{setspace}

%% EDITING MACROS
\usepackage{color}
\usepackage[section]{placeins}
\usepackage[normalem]{ulem}% to use \sout in feedback commands
\usepackage{bm}
% orange for EI
\definecolor{orange}{rgb}{1,0.5,0}
\newcommand\ei[2]{\sout{#1} \textcolor{orange}{#2}}
\newcommand\eic[1]{\textcolor{orange}{[#1]}}
% green for JW
\definecolor{green}{rgb}{0,0.5,0}
\newcommand\jw[2]{\sout{#1} \textcolor{green}{#2}}
\newcommand\jwc[1]{\textcolor{green}{[#1]}}
% purple for JJ
\definecolor{purple}{rgb}{0.5,0,1}
\newcommand\jj[2]{\sout{#1} \textcolor{purple}{#2}}
\newcommand\jjc[1]{\textcolor{purple}{[#1]}}
% cyan for AR
\definecolor{cyan}{rgb}{0,.5,.5}
\newcommand\ar[2]{\sout{#1} \textcolor{cyan}{#2}}
% \newcommand\arc[1]{\textcolor{cyan}{#1}}
% light brown for KT
\definecolor{lightbrown}{rgb}{0.5,0.5,0}
\newcommand\kt[2]{\sout{#1} \textcolor{lightbrown}{#2}}
\newcommand\ktc[1]{\textcolor{lightbrown}{#1}}

%% basic POMP definitions
\newcommand\Xspace{{\mathbb X}}
\newcommand\Yspace{{\mathbb Y}}
\newcommand\Thetaspace{\R^{\Thetadim}}
\newcommand\Xunitspace{{\scriptsize{\mathcal X}}}
\newcommand\Yunitspace{{\tiny{\mathcal Y}}}
\newcommand\hatTheta{\widehat{\Theta}}
\newcommand\Xdim{{\mathrm{dim}}(\Xspace)}
\newcommand\Ydim{{\mathrm{dim}}(\Yspace)}
\newcommand\Thetadim{P}
\newcommand\thetadim{p}
\newcommand\timeSet{{\mathbb T}}
\newcommand\data[1]{#1^*}

\newcommand\unitSpecific{\hspace{0.1mm}\mathrm{us}}
\newcommand\shared{\hspace{0.15mm}\mathrm{sh}}

% for algorithm pseudocode
\SetKwFor{For}{for}{:}{end}

%% for particle filters
\newcommand\unit{u}
\newcommand\altUnit{\tilde u}
\newcommand\Unit{U}
\newcommand\rep{i}
\newcommand\Rep{\mathcal{I}}
\newcommand\island{\rep}
\newcommand\Island{\Rep}
%\newcommand\txtisland{replicate}
\newcommand\altIsland{\tilde i}
\usepackage[mathscr]{euscript}

\renewcommand\time{n}
\newcommand\myvec[1]{\boldsymbol{#1}}
\newcommand\altTime{\tilde n}
\newcommand\Time{N}
\newcommand\Np{J}
\newcommand\np{j}
\newcommand\altNp{k}
\newcommand\altAltNp{\tilde j}
\newcommand\resampleIndex{r}
\newcommand\unitWeight{w}

%% for all iterated filtering algorithms
\newcommand\Nit{M}
\newcommand\nit{m}

%% customized math macros
\newcommand\prob{\mathbb{P}}
\newcommand\E{\mathbb{E}}
\newcommand\dd[1]{\mathrm{d}{#1}}
\newcommand\given{{\,\vert\,}}
\newcommand\equals{{{\,}={\,}}}
\newcommand\myequals{\hspace{0.5mm}{=}\hspace{0.5mm}}
\newcommand\myto{{\;:\;}}
\newcommand\seq[2]{{#1}\!:\!{#2}}
\newcommand\mydot{{\,\cdot\,}}
\newcommand\cp[2]{N_{\mathrm{#1}\mathrm{#2}}}
\newcommand\giventh{{\hspace{0.5mm};\hspace{0.5mm}}}
\newcommand\normal{\mathcal{N}}
\newcommand\argequals{{\,=\,}}
\newcommand\lags{c}
\newcommand\maxlag{\overline{c}}
\newcommand\nlfList{C}
\newcommand\bigO{\mathcal{O}}
\newcommand\loglik{\lambda}
\newcommand\loglikMC{\MC{\loglik}}
\newcommand\R{\mathbb{R}}
\newcommand\param{\,;}
\newcommand\mycolon{{\hspace{0.6mm}:\hspace{0.6mm}}}
\newcommand\MC[1]{#1^{\,\mbox{\tiny MC}}}
\newcommand\EMC{{\E}}
\newcommand\Var{\mathrm{Var}}
\newcommand\var{\Var}
\newcommand\Cov{\mathrm{Cov}}
\newcommand\cov{\Cov}
\newcommand\iid{\mathrm{iid}}
\newcommand\dist{d}
\newcommand\transpose{\top}

\newcommand\gnsep{,}

\def\lik{L}
\def\loglik{\ell}

%% THEOREM-LIKE ENVIRONMENTS
\usepackage{amsthm}
\newtheorem{prop}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{remark}{Remark}
\newtheorem{cor}{Corrolary}
\newtheorem{fact}{Fact}
\newtheorem{assumption}{Assumption}
\newtheorem{assumptionB}{Assumption}
\renewcommand\theassumption{A\arabic{assumption}}
\renewcommand\theassumptionB{B\arabic{assumptionB}}

\makeatletter
\AtBeginDocument{%
  \expandafter\renewcommand\expandafter\subsection\expandafter{%
    \expandafter\@fb@secFB\subsection
  }%
}
\makeatother

%% FOR PSEUDOCODE SETUP
\newcommand\mystretch{\rule[-2mm]{0mm}{5mm} }
\newcommand\asp{\hspace{4mm}}
\newcommand\codeIndent{\hspace{4mm}}

%%%%%%%%%%%%%%%% END OF USER DEFINITIONSS %%%%%%%%%%%%%%%

\bibliographystyle{apalike}
\usepackage{natbib}
\usepackage{fullpage}
% \usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
% \usepackage[labelfont=bf,labelsep=period]{caption} % For some reason this doesn't work with .Rnw file, but it does with just the tex file, so after compiling you can edit this so that it works.
\renewcommand{\figurename}{Fig}

\newcommand\myeqref[1]{(\ref{#1})}

\renewcommand{\contentsname}{Supplementary Content}
\renewcommand{\refname}{Supplementary References}
\renewcommand\thefigure{S-\arabic{figure}}
\renewcommand\thetable{S-\arabic{table}}
\renewcommand\thesection{S\arabic{section}}
\renewcommand\theequation{S\arabic{equation}}
\renewcommand\theprop{S\arabic{prop}}
\renewcommand\thelemma{S\arabic{lemma}}
\renewcommand\thealgocf{S\arabic{algocf}}

@@

\date{\today}
\title{Supplement to ``{\mytitle}''}
  \author{Jesse Wheeler, AnnaElaine L. Rosengart, Zhuoxun Jiang, \\ Kevin Tan, Noah Treutle and Edward L. Ionides
 \\ \hspace{.2cm}\\
    Department of Statistics, University of Michigan\\
}

\newcommand{\blind}{1}

\if1\blind
{
\maketitle
}
\fi

\if0\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf \mytitle}
\end{center}
  \bigskip
  \bigskip
}\fi

\tableofcontents

\newpage

\section{Model Diagrams}

Each of the dynamic models considered in this manuscript can be fully described using the model descriptions in the manuscript, coupled with the additional information described in Sections 2 and 3 of this supplement.
Despite this, diagrams of dynamic systems are often helpful to understand the equations.
In this section, we give three diagrams representing Models~1--3, respectively.
Because the models are defined by their mathematical equations and numeric implementation, these diagrams are not unique visual representations of the model.
Alternative representations that may be helpful in understanding the models explored in this paper were made by \citet{lee20sup}.

\subsection{Model~1}

\begin{figure}[!h]
%%%%% SEAIR diagram
\begin{center}
  \resizebox{0.9\textwidth}{!}{
    \Large
    \setlength{\unitlength}{5pt}
    \begin{picture}(100,95)(0,15)
      \thicklines

      % Unvaccinated Compartments
      \put(10,63){\framebox(6,6){$\mathrm{S}_0$}}
      \put(34,63){\framebox(6,6){$\mathrm{E}_0$}}
      \put(58,71){\framebox(6,6){$\mathrm{A}_0$}}
      \put(82,63){\framebox(6,6){$\mathrm{R}_0$}}
      \put(58,55){\framebox(6,6){$\mathrm{I}_0$}}

      % Births
      \put(1, 66){\vector(1, 0){8}}
      \put(4, 68){$\muBirth$}

      % S_0 -> E_0
      \put(17,66){\vector(1,0){16}}
      \put(22,67){$\lambda(t)$}

      % E_0 -> A_0
      \put(41,67.5){\vector(2.1,1){16}}
      % \put(37, 72){$\muEI\big(1-\symptomFrac_{\vaccCounter}(t)\big)$}

      % E_0 -> I_0
      \put(41,65.5){\vector(2.1,-1){16}}

      % A_0 -> R_0
      \put(65,74){\vector(2.1,-1){16}}

      % I_0 -> R_0
      \put(65,59){\vector(2.4,1){16}}

      % R_0 -> S_0
      \cbezier(84,70)(75, 95)(23,95)(14,70)
      \put(14,70){\vector(-1, -2){0.3}}

      % Unvaccinated death rates
      \put(13, 70){\vector(0, 1){10}}
      \put(11, 74){$\muDeath$}

      \put(37, 70){\vector(0, 1){10}}
      \put(35, 74){$\muDeath$}

      \put(61, 78){\vector(0, 1){9}}
      \put(59, 80){$\muDeath$}

      \put(61, 62){\vector(1.5, 2){6}}
      \put(61.5, 65.5){$\muDeath$}

      \put(85, 70){\vector(0, 1){10}}
      \put(83, 74){$\muDeath$}

      % Vaccinated Compartments
      \put(10,37){\framebox(6,6){$\mathrm{S}_\vaccCounter$}}
      \put(34,37){\framebox(6,6){$\mathrm{E}_\vaccCounter$}}
      \put(58,45){\framebox(6,6){$\mathrm{A}_\vaccCounter$}}
      \put(82,37){\framebox(6,6){$\mathrm{R}_\vaccCounter$}}
      \put(58,29){\framebox(6,6){$\mathrm{I}_\vaccCounter$}}

      % S -> E
      \put(17,40){\vector(1,0){16}}
      \put(22,41){$\lambda(t)$}

      % E -> A
      \put(41,40.5){\vector(2.1,1){16}}

      % E -> I
      \put(41,39.5){\vector(2.1,-1){16}}

      % A -> R
      \put(65,48){\vector(2.1,-1){16}}

      % I -> R
      \put(65,32){\vector(2.1,1){16}}

      % R -> S
      \cbezier(84, 36)(75, 10)(23, 10)(14, 36)
      \put(14,36){\vector(-1, 2){0.3}}

      % Vaccinated death rates
      \put(13, 36){\vector(0, -1){10}}
      \put(11, 31){$\muDeath$}

      \put(37, 36){\vector(0, -1){10}}
      \put(35, 31){$\muDeath$}

      \put(61, 28.5){\vector(0, -1){9}}
      \put(59, 24){$\muDeath$}

      \put(61, 44){\vector(1.5, -2){6}}
      \put(62, 38){$\muDeath$}

      \put(85, 36){\vector(0, -1){10}}
      \put(83.2, 30.2){$\muDeath$}

      % S_0 -> S_\vaccCounter
      \multiput(13,62)(0,-2){9}{\line(0,-1){1}}
      \put(13, 43.5){\vector(0, -1){0}}

      % E_0 -> E_\vaccCounter
      \multiput(37,62)(0,-2){9}{\line(0,-1){1}}
      \put(37, 43.5){\vector(0, -1){0}}

      % R_0 -> R_\vaccCounter
      \multiput(85,62)(0,-2){9}{\line(0,-1){1}}
      \put(85, 43.5){\vector(0, -1){0}}

      % A_0 -> A_\vaccCounter
      \multiput(57.5,72)(-1.9, 0){3}{\line(-1, 0){0.9}}
      \put(52.2, 72){\line(-2.1, -1){1}}
      \multiput(51,71)(0, -2){11}{\line(0, -1){1}}
      \put(52.2, 49){\line(-2.1, 1){1}}
      \multiput(54,49)(1.9, 0){2}{\line(-1, 0){0.9}}
      \put(57, 49){\vector(1, 0){0}}

      % I_0 -> I_\vaccCounter
      \multiput(65, 57.5)(1.9, 0){3}{\line(1, 0){0.9}}
      \put(70.6, 57.5){\line(2.1, -1){1}}
      \multiput(72, 56.5)(0, -2){13}{\line(0, -1){1}}
      \put(72, 30.9){\line(-2, -1.2){0.9}}
      \multiput(65.6, 30.2)(1.9, 0){3}{\line(1, 0){0.9}}
      \put(64.5, 30.2){\vector(-1, 0){0}}

    \end{picture}
  }
\end{center}
\caption{A flow diagram for the SEAIR model.}\label{fig:flow_diagram}
\end{figure}

\subsection{Model~2}

\begin{figure}[!h]
%%%%% SEAIR diagram
\begin{center}
  \resizebox{\textwidth}{!}{
    \Large
    \setlength{\unitlength}{5pt}
    \begin{picture}(100,95)(0,15)
      \thicklines

      % Unvaccinated Compartments
      \put(10,63){\framebox(6,6){$\mathrm{S}_{u0}$}}
      \put(34,69){\framebox(6,6){$\mathrm{E}_{u0}$}}
      \put(58,77){\framebox(6,6){$\mathrm{A}_{u0}$}}
      \put(82,77){\framebox(6,6){$\mathrm{R^A}_{u0}$}}
      \put(82,61){\framebox(6,6){$\mathrm{R^I}_{u0}$}}
      \put(58,61){\framebox(6,6){$\mathrm{I}_{u0}$}}

      % % S_0 -> E_0
      \put(17,66){\vector(3,1.1){16}}
      % \put(22,67){$\lambda(t)$}

      % % E_0 -> A_0
      \put(41,72.5){\vector(2.1,1){16}}
      % % \put(37, 72){$\muEI\big(1-\symptomFrac_{\vaccCounter}(t)\big)$}

      % % E_0 -> I_0
      \put(41,71.5){\vector(2.1,-1){16}}

      % % A_0 -> R_0
      \put(65,80.5){\vector(1,0){16}}

      % A_0 -> W_u
      % \multiput(65, 79.8)(1, -1){8}{\line(1, -1){0.25}}

      % I_0 -> W_u
      % \multiput(65, 64.2)(1, 1){8}{\line(1, 1){0.25}}

      % Long verticle dotted line
      % \multiput(72.9, 72)(0, -1.41){13}{\line(0, -1){0.25}}

      % Long backwards horizontal dotted line
      % \multiput(73, 54)(-1.41, 0){22}{\line(0, -1){0.25}}
      % \multiput(73, 34)(0, 1.41){15}{\line(0, 1){0.25}}
      % \put(42, 53.9){\vector(-1, 0){0.2}}

      % A_uz -> W_u
      \put(58,39){\framebox(6,6){$\mathrm{A}_{u\vaccCounter}$}}
      \qbezier[25](60.5, 46)(51, 52)(42, 53)
      \put(42, 53){\vector(-1, 0.1){0.5}}

      % A_uz -> W_u
      \qbezier[25](60.5, 60)(51, 54)(42, 54)
      \put(42, 54){\vector(-1, -0.1){0.5}}

      % I_uz -> W_u
      \qbezier[30](60.5,30)(48, 45)(40, 50)
      \put(40, 50){\vector(-1, 0.8){0.5}}

      % A_u0 -> W_u
      \qbezier[29](60.5,76)(48, 61)(40.6, 57)
      \put(40.8, 57.1){\vector(-1, -0.7){0.5}}

      % A + I_\vaccCounter -> W_u
      % \multiput(65, 41.8)(1, -1){8}{\line(1, -1){0.25}}
      % \multiput(65, 26.2)(1, 1){8}{\line(1, 1){0.25}}

      % W_u -> backwards diags.
      % \multiput(33, 57)(-1, 1.2){10}{\line(-1, 1.2){0.1}}
      % \put(24.5, 67){\vector(-1, 1.2){0.5}}
      % \multiput(33, 51)(-1, -1.2){10}{\line(-1, -1.2){0.1}}
      % \put(24, 40){\vector(-1, -1.2){0.5}}
      %\put(65, 80){\vector(2, -1){8}}

      % % I_0 -> R_0
      \put(65,63.5){\vector(1,0){16}}

      % % Vaccinated Compartments
      \put(10,37){\framebox(6,6){$\mathrm{S}_{u\vaccCounter}$}}
      \put(34,31){\framebox(6,6){$\mathrm{E}_{u\vaccCounter}$}}
      \put(58,39){\framebox(6,6){$\mathrm{A}_{u\vaccCounter}$}}
      \put(82,39){\framebox(6,6){$\mathrm{R^A}_{u\vaccCounter}$}}
      \put(58,23){\framebox(6,6){$\mathrm{I}_{u\vaccCounter}$}}
      \put(82,23){\framebox(6,6){$\mathrm{R^I}_{u\vaccCounter}$}}

      % R_0 -> S_0
      \cbezier(85,22)(80, 8)(18,8)(12,36)
      \put(12.1,35.5){\vector(-1, 3.5){0.3}}

      \cbezier(85,38)(80, 12)(18,15)(14,36)
      \put(14.1,35.5){\vector(-1, 3.5){0.3}}

      % R_0 -> S_0
      \cbezier(85,84)(80, 100)(18,100)(12,70)
      \put(12.1,70.5){\vector(-1, -3.5){0.3}}

      \cbezier(85,68)(80, 95)(18, 92)(14,70)
      \put(14.1,70.5){\vector(-1, -3.5){0.3}}

      % % S -> E
      \put(17,40){\vector(3,-1.1){16}}
      % \put(22,41){$\lambda(t)$}

      % % E -> A
      \put(41,34.5){\vector(2.1,1){16}}

      % % E -> I
      \put(41,33.5){\vector(2.1,-1){16}}

      % % A -> R
      \put(65,42.5){\vector(1,0){16}}

      % % I -> R
      \put(65,25.5){\vector(1,0){16}}

      % S_0 -> S_\vaccCounter
      \multiput(12,62)(0,-2){9}{\line(0,-1){1}}
      \put(12, 44){\vector(0, -1){0}}

      % S_0 -> S_\vaccCounter
      % Wanning vaccine efficacy
      \multiput(14,44)(0,2){9}{\line(0,1){1}}
      \put(14, 62){\vector(0, 1){0}}

      \put(37, 54){\circle{7}}
      \put(35, 53){$\mathrm{W_u}$}

      \qbezier[10](32.5, 54)(29.5, 54)(26.5, 54)
      \put(26, 54){\vector(-1, 0){0.1}}
      \put(28, 55){$\Wremoval$}

    \end{picture}
  }
\end{center}
\caption{A flow diagram for the SEAIR model 2. This is a constant population model, there are no births/deaths. Vaccinations are assumed to only be given to susceptible individuals, and vaccine immunity wanes only with susceptible vaccinated individuals.}\label{fig:flow_diagram2}
\end{figure}

\newpage

\subsection{Model~3}

\begin{figure}[!h]
%%%%% SEAIR diagram
\begin{center}
  \resizebox{\textwidth}{!}{
    \Large
    \setlength{\unitlength}{5pt}
    \begin{picture}(100,85)(0,15)

    % COMPARTMENTS
    % \put(36, 47){\framebox(6, 6){$\mathrm{W}_{u}$}}

    \put(39, 50){\circle{6}}
    \put(37, 49){$\mathrm{W}_{u}$}

    \put(8, 55){\framebox(6, 6){$\mathrm{S_{u0}}$}}
    \put(8, 39){\framebox(6, 6){$\mathrm{S_{u\vaccCounter}}$}}
    % \put(22, 45){\circle{2}}
    % \put(22, 55){\circle{2}}
    \put(36, 62.5){\framebox(6, 6){$\mathrm{A_{u}}$}}
    \put(36, 31.5){\framebox(6, 6){$\mathrm{I_{u}}$}}
    \put(56, 55){\framebox(6, 6){$\mathrm{R^1_{u0}}$}}
    \put(71, 55){\framebox(6, 6){$\mathrm{R^2_{u0}}$}}
    \put(86, 55){\framebox(6, 6){$\mathrm{R^3_{u0}}$}}
    \put(56, 39){\framebox(6, 6){$\mathrm{R^1_{u\vaccCounter}}$}}
    \put(71, 39){\framebox(6, 6){$\mathrm{R^2_{u\vaccCounter}}$}}
    \put(86, 39){\framebox(6, 6){$\mathrm{R^3_{u\vaccCounter}}$}}
    % \dottedline{3}(0,5)(70,5)

    % INDIVIDUAL MOVEMENT
    % \put(15, 58){\line(2.4, -1){5.5}}
    % \put(15, 42){\line(2.4, 1){5.5}}

    % \put(23, 46.3){\vector(1, 1.45){12}}
    % \put(23, 53.7){\vector(1, -1.45){12}}

    % \put(23.5, 55.7){\vector(1.1, 1){11}}
    % \put(23.5, 44.3){\vector(1.1, -1){11}}

    % Backward Diags, W_u
    % \multiput(35, 51)(-1, .3){12}{\line(-1, .3){0.2}}
    % \put(24, 54.4){\vector(-1, 0.3){0.5}}
    % \multiput(35, 49)(-1, -0.3){12}{\line(-1, -0.3){0.15}}
    % \put(24, 45.6){\vector(-1, -0.3){0.5}}

    % S_u0 -> A_u
    \put(15, 58.5){\vector(3, 1.2){20}}

    % S_u0 -> I_u
    \put(15, 57.5){\vector(1, -1.1){20}}

    % S_uz -> I_u
    \put(15, 41.5){\vector(3, -1.2){20}}

    % S_uz -> A_u
    \put(15, 42.5){\vector(1, 1.1){20}}

    % I_u -> R_u0
    \put(43, 66){\vector(2, -1){12}}

    % A_u -> R_uz
    \multiput(43, 64)(1.25, -1.6){12}{\line(1, -1.3){0.7}}
    \put(57, 46){\vector(1, -1.3){0.2}}

    \put(43, 36){\vector(1.7, 2){15}}

    \multiput(60, 54)(0, -2){4}{\line(0, -1){1}}
    \put(60, 46){\vector(0, -1){0.5}}

    \multiput(74, 54)(0, -2){4}{\line(0, -1){1}}
    \put(74, 46){\vector(0, -1){0.5}}

    \multiput(89, 54)(0, -2){4}{\line(0, -1){1}}
    \put(89, 46){\vector(0, -1){0.5}}

    \put(63, 58){\vector(1, 0){7}}
    \put(78, 58){\vector(1, 0){7}}

    \put(63, 42){\vector(1, 0){7}}
    \put(78, 42){\vector(1, 0){7}}

    \multiput(39, 61.5)(0, -1){8}{\line(0, -1){0.3}}
    \put(39,54){\vector(0, -1){0.5}}

    \multiput(39, 38.5)(0, 1){8}{\line(0, 1){0.3}}
    \put(39,46){\vector(0, 1){0.5}}

    \cbezier(89,62)(80, 85)(20, 85)(12,62)
    \put(12.2,62.5){\vector(-1, -3.5){0.3}}

    \cbezier(89,38)(80, 15)(20, 15)(12,38)
    \put(12.2,37.5){\vector(-1, 3.5){0.3}}

    \multiput(11, 54)(0, -2){4}{\line(0, -1){1}}
    \put(11, 46){\vector(0, -1){0.5}}

    % DEMOGRAPHY
    % -> S
    \put(0, 58){\vector(1, 0){7}}
    \put(1, 59){$\mathrm{\mu_{\demography S_{u0}}}$}

    % S Death
    \put(10, 62){\vector(0, 1){7}}
    \put(8, 64){$\mathrm{\muDeath}$}
    \put(10, 38){\vector(0, -1){7}}
    \put(8, 34){$\mathrm{\muDeath}$}

    % A + I death
    \put(39, 69.5){\vector(0, 1){7}}
    \put(37, 71.5){$\mathrm{\muDeath}$}

    \put(39, 31.5){\vector(0, -1){7}}
    \put(30, 27.5){$\mathrm{\muDeath + \choleraDeath}$}

    % R Death
    \put(59, 62){\vector(0, 1){7}}
    \put(57, 64){$\mathrm{\muDeath}$}

    \put(74, 62){\vector(0, 1){7}}
    \put(72, 64){$\mathrm{\muDeath}$}

    \put(59, 38){\vector(0, -1){7}}
    \put(57, 34){$\mathrm{\muDeath}$}

    \put(74, 38){\vector(0, -1){7}}
    \put(72, 34){$\mathrm{\muDeath}$}

    \put(93, 42){\vector(1, 0){7}}
    \put(95.5, 43){$\mathrm{\muDeath}$}

    \put(93, 58){\vector(1, 0){7}}
    \put(95.5, 59){$\mathrm{\muDeath}$}

    % W death
    % \put(43, 50){\vector(1, 0){7}}
    \multiput(43, 50)(1, 0){7}{\line(1, 0){0.2}}
    \put(50.2, 50){\vector(1, 0){0.5}}
    \put(44, 51){$\mathrm{\Wremoval}$}
    % \qbezier[12](10, 20)(20, 30)(25, 20)

    \end{picture}
  }
\end{center}
\caption{A flow diagram for the SAIR model 3. This model assumes a constant population while also including a mechanism for births/deaths; all deaths are balanced by births into the unvaccinated susceptible compartment, so the birth rate $\mu_{\demography S_{u0}}$ corresponds to the sum total deaths from the remaining compartments. The model assumes that symptomatic individuals will not be vaccinated, hence no vaccination arrow exiting the $I_u0$ compartment.}\label{fig:flow_diagram3}
\end{figure}

\section{Markov chain and differential equation interpretations of compartment flow rates}

In Sections 2.1, 2.2 and 2.3 of the main article, we define compartment models in terms of their flow rates.
For a discrete population model, these rates define a Markov chain.
For a continuous and deterministic model, the rates define a system of ordinary differential equations.
Here, we add additional details to clarify the mapping from a collection of rate functions to a fully specified process.
Our treatment follows \citet{breto09}.

A general compartment model is a vector-valued process $X(t)=(X_1(t),\dots,X_c(t))$ denoting the (integer or real-valued) counts in each of $c$ compartments, where $t$ is any continuous value in the interval  $\left[t_0, \infty\right)$ for some real valued starting time $t_0$.
The compartments may also have names, but to set up general notation we simply refer to them by their numerical index.
The basic characteristic of a compartment model is that $X(t)$ can be written in terms of the flows $N_{ij}(t)$ from $i$ to $j$.
A flow into compartment $i$ from outside the system is denoted by $N_{\demography i}$, and a flow out of the system from compartment $i$ is denoted by $N_{i\demography}$.
We call $\demography$ a source/sink compartment, though it is an irregular compartment since $X_{\demography}(t)$ is not defined.
These flows are required to satisfy a ``conservation of mass'' identity:
\begin{equation}
X_i(t)=X_i(t_0)+N_{\demography i}(t) - N_{i\demography}(t) + \sum_{j\neq
i}N_{ji}(t)-\sum_{j\neq i}N_{ij}(t). \label{eq:conservation}
\end{equation}
Each {\em flow} $N_{ij}(t)$ is associated with a {\em rate} function
$\mu_{ij}=\mu_{ij}(t,X(t))$, where we include the possibility that $i$ or $j$ takes value $\demography$.

There are different ways to use a collection of rate functions to build a fully specified model.
We proceed to describe the ones we use in this paper: via a system of ordinary differential equations (Sec.~\ref{subsec:ode}), a simple Markov counting system (Sec.~\ref{subsec:smcs}), and an overdispersed Markov counting system (Sec.~\ref{subsec:odmcs}). Other representations include stochastic differential equations driven by Gaussian noise or Gamma noise \citep{bhadra11}.

\subsection{Ordinary differential equation (ODE) interpretation}
\label{subsec:ode}

A basic deterministic specification is
\begin{equation}
\label{eq:ode1}
dN_{ij}/dt = \mu_{ij}\big(t,X(t) \big) X_i(t), \hspace{3mm} i\in \seq{1}{c}, \hspace{3mm} j\in \seq{1}{c} \cup\{\demography\}, \hspace{3mm} i\neq j,
\end{equation}
where $\mu_{ij}\big(t,X(t)\big)$ is called a per-capita rate or a unit rate.
Flows into the system require special treatment since $X_i(t)$ in \myeqref{eq:ode1} is not defined for $i=\demography$.
Instead, we specify
\begin{equation}
\label{eq:ode2}
dN_{\demography i}/dt = \mu_{\demography i}\big(t,X(t) \big).
\end{equation}
This is the the interpretation and implementation used for Model~2 in our study.

>>>>

\input{inputs/header}

<<packages, include=FALSE>>=
library(tidyverse)
library(pomp)
library(haitipkg)
library(spatPomp)
library(doParallel)
library(doRNG)
library(knitr)

options(
  scipen = 2,
  help_type = "html",
  stringsAsFactors = FALSE,
  continue = "+  ",
  width = 70,
  useFancyQuotes = FALSE,
  reindent.spaces = 2,
  xtable.comment = FALSE
)

opts_chunk$set(
  tidy.opts=list(
    keep.blank.line=FALSE
  ),
  comment="",
  fig.path = 'figure/',
  # warning=FALSE,
  # message=FALSE,
  echo=FALSE,
  fig.align="center",
  dev="pdf",
  dev.args=list(
    bg="transparent",
    pointsize=9
  )
)

@@

@@

%%%%%%%%%% START
